---
title: "SQL"
output: html_document
---
 
## FILTERING COLUMNS AND ROWS
In SQL one can filter columns and rows of a selected table. Columns can be filtered using the SELECT command and rows can be filtered using the WHERE command. Following is an example of this using the Lahman baseball database.

Before the tables can be filtered, the Lahman and sqldf packages need to be loaded.
One can become more selective with the Batting data, using the same filters.  

### Examples 
Question1: Suppose we wanted to find all the Yankees that hit greater than 40 homeruns.  We could write the following code:


```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"

sqldf(query)
```


Question2: Suppose we wanted to find all the stances where a player had more than 40 homeruns but less than 60 strikeouts. We would do this in the following way:

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)

query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"

sqldf(query)

```

Question3: Suppose we wanted to find an instances of Phillies in the the 1970s hitting more than 30 Homeruns. We would do this in the following way:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID>=1970 and yearID<=1980"

sqldf(query)


```

## ORDERING DATA OUTPUTS
With SQL, you can also order the rows in the output. To order the information, an ORDER BY clause is used. This will default to ascending order (smallest to largest). To organize in descending order, the DESC command will be placed after the ORDER BY command.

### Examples 

Question1: Suppose you wanted to see evrey instance of a player hitting more than 50 homeruns.  But you would like the players with the most homeruns to be at the top of this list.  We would do this in the following way:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```


Suppose we wanted to find all instances of a player striking out less than 10 times, with at least 400 at-bats. We would do this in the following code:

```{r}
library(Lahman)
library(sqldf)

query<-"SELECT playerID,teamID,yearID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"

sqldf(query)

```


## AGGREGATION 
Aggregation is essentially the "squishing" of all the data into 1 record. It will output the sum of all records. This is done by using the GROUP BY command. When using GROUP BY, one must then specify how the other columns asked for should be grouped. If one wants to continue filtering after they have grouped, they can use the HAVING command, which functions as the filtering WHERE command when used after GROUP BY. ORDER BY can be used in the same way as before, coming at the end of the command list.

### Examples 
Question1: Suppose we wanted to find the career homerun totals for Babe Ruth.  We would do this using the following code:

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)
```

Question2: Suppose we wanted to find career HR totals for all players, but list only those with total of 600 or more. Also we wanted to order them by having greatest total at the top.  We would do this using the following code: 

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

Question3: Suppose we wanted to find what players have the highest average per season homeruns totals and we wanted to list only the average greater than 30 homeruns.  We would do this using the following code:

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>=30
ORDER BY avg(HR) DESC"

sqldf(query)
```

## JOINS
Joining allows tables to be joined together based on common data to increase the data output options. This is done through the INNER JOIN and ON commands. INNER JOIN indicates which tables are being joined together. ON indicates the common data that the tables will be joined by.

### Examples 
Question1: Suppose we wanted to find the first and last names along with team, year, and homeruns for Babe Ruth. We would do this using the following code: 

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Question2: Suppose we wanted to find all instances of players hitting more than 50 homeruns and incude first and last names, team, year, and homeruns.  We would do this using the following code:

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.HR>50
ORDER BY HR DESC"

sqldf(query)
```

Question3: Suppose we wanted to find Babe Ruth's homeruns by season and what team. We would do this using the following code: 

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

